---
title: "hk_1_0918"
author: "Ruoyuan Qian"
date: "2019/9/18"
output: pdf_document
---

# Problem 1

## a)

I found that there are two rows of variable names which may cause some errors when reading the data into R. As a result, I use `skip=` to skip first row and read others into R.

```{r Poblem1 a read data, message = FALSE,warning = FALSE,echo = FALSE}

# make sure to install "formattable" package first!!!!!!!!!!!!!!!!

library(formattable)
library(readxl)
library(dplyr)
library(arsenal)
library(ggplot2)

 raw_data<-read_excel(".\\Exercise.xlsx",skip =1)

#check variables and values in raw_data
 str(raw_data)                
 
#check if there is a missing value
 anyNA(raw_data) 
 
#only 3 kinds of race in the data
 table(raw_data$Race)         
```
```{r Poblem1 a summary data,echo = FALSE}


#rename some values into a readable way 
 fixed_data    <-     raw_data %>% mutate(
         Group  =     recode(Group,  "0" = "Control","1" = "Intervention"),
         Gender =     recode(Gender, "1" = "male", "2" = "female"),
         HTN    =     recode(HTN,    "0" = "no",   "1" = "yes"),
         T2DM   =     recode(T2DM,   "0" = "no",   "1" = "yes"),
         Smokes =     recode(Smokes, "0" = "no",   "1" = "yes"),
         Depression = recode(Depression, "0" = "no", "1" = "yes"),
         Race   =     recode(Race,   "1" = "African American", 
                                     "2" = "Hispanic",
                                     "3" = "African American",
                                     "4" = "G"))
#rename variables
 names(fixed_data)[c(9:20)]=c("SYS_PRE","SYS_POST","DIA_PRE","DIA_POST",
                              "BMI_PRE","BMI_POST","HDL_PRE","HDL_POST",
                              "LDL_PRE","LDL_POST","GLU_PRE","GLU_POST")

#draw the descriptive table
 sum_data  <-  arsenal::tableby(Group ~ Age + Gender + Race +
                                        Depression + Smokes + HTN+T2DM, 
                                data  = fixed_data,
                                test  = FALSE, 
                                total = FALSE,
                                numeric.stats = c("meansd" ,"median" ) )
 summary(sum_data,text = TRUE)

```
## b)

### b) i
```{r b) i,echo = FALSE}

#install.packages("formattable")
# seperately select data from Intervention and Control
# a denotes data from Intervention
  data_1 <- filter(fixed_data,Group=="Intervention")
  a      <- as.matrix(data_1[,c(9:20)])

# b denotes data from Control
  data_2 <- filter(fixed_data,Group=="Control")
  b      <- as.matrix(data_2[,c(9:20)])
  
# use a loop to calculate 4 coluumns of values:  baseline and six_month in control and Intervention, respectively
## define four columns
  baseline_inter  <- NULL
  six_month_inter <- NULL
  baseline_con    <- NULL
  six_month_con   <- NULL
  
for(i in 1:6){ 
  
  tem_mean    <- digits(meansd(a[,2*i-1]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  pre_median  <- median  (a[,2*i-1])
  tem_quan    <- quantile(a[,2*i-1])
  tem_medran  <- paste   (pre_median ,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  tem_delta   <- digits(meansd(a[,2*i]-a[,2*i-1]),2)
  tem_delsd   <- paste (tem_delta[1],"±",tem_delta[2])

  baseline_inter <- rbind(baseline_inter, tem_meansd, 
                              tem_medran, tem_delsd)
  
  tem_mean    <- digits(meansd(b[,2*i-1]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  pre_median  <- median  (b[,2*i-1])
  tem_quan    <- quantile(b[,2*i-1])
  tem_medran  <- paste   (pre_median ,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  tem_delta   <- digits(meansd(b[,2*i]-b[,2*i-1]),2)
  tem_delsd   <- paste (tem_delta[1],"±",tem_delta[2])
  
  baseline_con   <- rbind(baseline_con, tem_meansd, 
                            tem_medran,tem_delsd)

    
  tem_mean    <- digits(meansd(a[,2*i]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  POST_median <- median  (a[,2*i])
  tem_quan    <- quantile(a[,2*i])
  tem_medran  <- paste   (POST_median,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  
  six_month_inter <- rbind(six_month_inter, tem_meansd, 
                                tem_medran," ")
  
  tem_mean    <- digits(meansd(b[,2*i]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  POST_median <- median  (b[,2*i])
  tem_quan    <- quantile(b[,2*i])
  tem_medran  <- paste   (POST_median,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  six_month_con  <- rbind(six_month_con, tem_meansd, 
                             tem_medran," ")
}
  
# Add row names in a column
  row_name <- c("SYS","DIA","BMI","HDL","LDL","GLU")
  rname    <- NULL
  
  for(i in row_name){
   rname   <- rbind(rname,i," ","delta")
                    }

# combine all columns into a dataframe with more readable names
  df <- tibble("Variable"          = rname,
               "Intervention,N=36" = baseline_inter,
               "Intervention,N=36 "= six_month_inter,
               "Control,N=36"      = baseline_con,
               "Control,N=36 "     = six_month_con)

  df_compare <- rbind(c(" ","baseline","six month",
                            "baseline","six month",
                            "baseline","six month"),df)

# save the data frame as a .csv file 
 write.csv(df_compare, file="df_compare.csv")

# read the .csv tocheck the result
 read.csv ("df_compare.csv")

```

### b) ii
```{r b) ii,echo = FALSE}

# merge BMI_PRE and BMI_POST into one column
  a <- dplyr::select(fixed_data, c(Group,BMI_PRE ))
  b <- dplyr::select(fixed_data, c(Group,BMI_POST))
  bmi_pre  <- rename(a,BMI=BMI_PRE )
  bmi_post <- rename(b,BMI=BMI_POST)
  
# add a new variable that helps to illustrate time status of BMI, either baseline or  post-measurement
  len_pre  <- length( bmi_pre$Group )
  len_post <- length( bmi_post$Group )
  data_b   <- rbind ( cbind ( bmi_pre, 
                              time=rep(0,len_pre )),
                      cbind ( bmi_post,
                              time=rep(1,len_post)))
  data_bmi <- data_b %>% mutate(
              time = recode(time, "0" = "baseline", 
                                  "1" = "post-measurement"))
  

# merge DLD_PRE and DLD_POST into one column
 a1 <- dplyr::select(fixed_data, c(Group,LDL_PRE ))
 b1 <- dplyr::select(fixed_data, c(Group,LDL_POST))
 ldl_pre  <- rename ( a1,LDL = LDL_PRE )
 ldl_post <- rename ( b1,LDL = LDL_POST)
 
# add a new variable that helps to illustrate time status of DLD, either baseline or  post-measurement
 len_pre  <- length( ldl_pre$Group  )
 len_post <- length( ldl_pre$Group  )
 data_l   <- rbind ( cbind(ldl_pre,time=rep(0,len_pre)),
                     cbind(ldl_post,time=rep(1,len_post)))
 data_ldl <- data_l %>% mutate(
             time = recode(time, "0" = "baseline", 
                                 "1" = "post-measurement"))

```

```{r ii ,echo = FALSE}
  B<-tibble(  group  = pull(data_bmi,Group),
              bmi    = pull(data_bmi,BMI) ,
              time   = pull(data_bmi,time)) 
  L<-tibble(  group  = pull(data_ldl,Group),
              ldl    = pull(data_ldl,LDL) ,
              time   = pull(data_ldl,time)) 
 
   ggplot(B,aes(x    = group, 
                       y    = bmi, 
                       fill = time)) +
   geom_boxplot() +
   scale_fill_manual(values = c("blue", "red"), 
                     labels = expression("baseline", 
                                         "post-measurement")) + 
   xlab("") + 
   ylab("BMI ") + 
   theme(legend.position = "top", legend.title = element_blank())
   
   
    
   ggplot(L,aes(x    = group, 
                       y    = ldl, 
                       fill = time)) +
   geom_boxplot() +
   scale_fill_manual(values = c("blue", "red"), 
                     labels = expression("baseline", 
                                         "post-measurement")) + 
   xlab("") + 
   ylab("BMI ") + 
   theme(legend.position = "top", legend.title = element_blank())
```


### b) iii



According to table of Pre/Post changes in metabolic parameters, structured exercise program has a greater impact on the`Systolic`, `Glu`  variables in the intervention group than on the corresponding variables in the control group. So it means that the program may affect `Systolic` , `Glu` and `BMI`more apparently compared with the impact on the other variables. 

Moreover, there are some variables that changed more severely in control than in intervention group, such as `diastolic`, `HDL` and `LDL`. In other words, the program is not benefial to the decline of `diastolic`, `HDL` and `LDL`.

As for the BMI in the study, in general, the values in intervention group are higher than the control group's while there is little difference of values within each group. 

As for the LDL, for both groups, there is a trend that the values in post measurement are lower than the baseline's, and it is more obvious in control. Besides, in intervention , the values are more concentrated whereas values in control are more dispersed.



## c) 

It is an interventional clinical trial, and from the perspective of demographics, the two groups are basically balanced, which is favorable for the study.

However, there is some potential issues as well. For instance, the medians of baselines of BMI in two groups are of great difference, which may make the two groups incomparable. Besides, as for the LDL, the variances of two groups are extremely different, which may cause the similar problem like BMI. 

Besides, the study dose not implement binding, so there may contain a lot of bias from doctors, subjects and so on. What is worse,the control is non-participating, as a result, the new program in the study is just compared with blank control instead of standard control. Due to that, the study can only tell the difference btween new program and non-participant but fails to test the difference btween new program and established one. 

To sum up, the study contains plenty of bias and the result is not convincing enough.

# Problem 2

The probability is The probability is`(0.6×0.001)/(0.6×0.001+0.05×0.999)=0.012`,

# Problem 3

aaa





