---
title: "p8130_hw1_rq2166"
author: "Ruoyuan Qian"
date: "2019/9/20"
output:   
  html_document:
    toc: true
    toc_float: true
---

# Problem 1

## a)

I found that there are two rows of variable names which may cause some errors when reading the data into R. As a result, I use `skip=` to skip first row and read others into R.

!!!Please notice that install package `"formattable"` first if your computer dosen't have it.!!!

read data and check its basic information
```{r Poblem1 a read data, message = FALSE,warning = FALSE}

# make sure to install "formattable" package first!!!!!!!!!!!!!!!!

library(formattable)
library(readxl)
library(dplyr)
library(arsenal)
library(ggplot2)

 raw_data<-read_excel(".\\Exercise.xlsx",skip =1)

#check variables and values in raw_data
 str(raw_data)                
 
#check if there is a missing value
 anyNA(raw_data) 
 
#only 3 kinds of race in the data
 table(raw_data$Race)         
```

Create table of demographics and co-morbidities 
```{r Poblem1 a summary data}


#rename some values into a readable way 
 fixed_data    <-     raw_data %>% mutate(
         Group  =     recode(Group,  "0" = "Control","1" = "Intervention"),
         Gender =     recode(Gender, "1" = "male", "2" = "female"),
         HTN    =     recode(HTN,    "0" = "no",   "1" = "yes"),
         T2DM   =     recode(T2DM,   "0" = "no",   "1" = "yes"),
         Smokes =     recode(Smokes, "0" = "no",   "1" = "yes"),
         Depression = recode(Depression, "0" = "no", "1" = "yes"),
         Race   =     recode(Race,   "1" = "African American", 
                                     "2" = "Hispanic",
                                     "3" = "African American",
                                     "4" = "G"))
#rename variables
 names(fixed_data)[c(9:20)]=c("SYS_PRE","SYS_POST","DIA_PRE","DIA_POST",
                              "BMI_PRE","BMI_POST","HDL_PRE","HDL_POST",
                              "LDL_PRE","LDL_POST","GLU_PRE","GLU_POST")

#draw the descriptive table
 sum_data  <-  arsenal::tableby(Group ~ Age + Gender + Race +
                                        Depression + Smokes + HTN+T2DM, 
                                data  = fixed_data,
                                test  = FALSE, 
                                total = FALSE,
                                numeric.stats = c("meansd" ,"median" ) )
 summary(sum_data,text = TRUE)

```
## b)

### b) i

create table of pre/post changes in metabolic parameters
```{r b) i}

#install.packages("formattable")
# seperately select data from Intervention and Control
# a denotes data from Intervention
  data_1 <- filter(fixed_data,Group=="Intervention")
  a      <- as.matrix(data_1[,c(9:20)])

# b denotes data from Control
  data_2 <- filter(fixed_data,Group=="Control")
  b      <- as.matrix(data_2[,c(9:20)])
  
# use a loop to calculate 4 coluumns of values:  baseline and six_month in control and Intervention, respectively
## define four columns
  baseline_inter  <- NULL
  six_month_inter <- NULL
  baseline_con    <- NULL
  six_month_con   <- NULL
  
for(i in 1:6){ 
  
  tem_mean    <- digits(meansd(a[,2*i-1]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  pre_median  <- median  (a[,2*i-1])
  tem_quan    <- quantile(a[,2*i-1])
  tem_medran  <- paste   (pre_median ,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  tem_delta   <- digits(meansd(a[,2*i]-a[,2*i-1]),2)
  tem_delsd   <- paste (tem_delta[1],"±",tem_delta[2])

  baseline_inter <- rbind(baseline_inter, tem_meansd, 
                              tem_medran, tem_delsd)
  
  tem_mean    <- digits(meansd(b[,2*i-1]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  pre_median  <- median  (b[,2*i-1])
  tem_quan    <- quantile(b[,2*i-1])
  tem_medran  <- paste   (pre_median ,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  tem_delta   <- digits(meansd(b[,2*i]-b[,2*i-1]),2)
  tem_delsd   <- paste (tem_delta[1],"±",tem_delta[2])
  
  baseline_con   <- rbind(baseline_con, tem_meansd, 
                            tem_medran,tem_delsd)

    
  tem_mean    <- digits(meansd(a[,2*i]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  POST_median <- median  (a[,2*i])
  tem_quan    <- quantile(a[,2*i])
  tem_medran  <- paste   (POST_median,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  
  six_month_inter <- rbind(six_month_inter, tem_meansd, 
                                tem_medran," ")
  
  tem_mean    <- digits(meansd(b[,2*i]),2)
  tem_meansd  <- paste (tem_mean[1],"±",tem_mean[2])
  
  POST_median <- median  (b[,2*i])
  tem_quan    <- quantile(b[,2*i])
  tem_medran  <- paste   (POST_median,"(",
                          tem_quan[2],"-",
                          tem_quan[4],")")
  
  six_month_con  <- rbind(six_month_con, tem_meansd, 
                             tem_medran," ")
}
  
# Add row names in a column
  row_name <- c("SYS","DIA","BMI","HDL","LDL","GLU")
  rname    <- NULL
  
  for(i in row_name){
   rname   <- rbind(rname,i," ","delta")
                    }

# combine all columns into a dataframe with more readable names
  df <- tibble("Variable"          = rname,
               "Intervention,N=36" = baseline_inter,
               "Intervention,N=36 "= six_month_inter,
               "Control,N=36"      = baseline_con,
               "Control,N=36 "     = six_month_con)

  df_compare <- rbind(c(" ","baseline","six month",
                            "baseline","six month",
                            "baseline","six month"),df)

# save the data frame as a .csv file 
 write.csv(df_compare, file="df_compare.csv")

# read the .csv tocheck the result
 read.csv ("df_compare.csv")

```

### b) ii

draw plots of BMI and LDL
```{r b) ii}

# merge BMI_PRE and BMI_POST into one column
  a <- dplyr::select(fixed_data, c(Group,BMI_PRE ))
  b <- dplyr::select(fixed_data, c(Group,BMI_POST))
  bmi_pre  <- rename(a,BMI=BMI_PRE )
  bmi_post <- rename(b,BMI=BMI_POST)
  
# add a new variable that helps to illustrate time status of BMI, either baseline or  post-measurement
  len_pre  <- length( bmi_pre$Group )
  len_post <- length( bmi_post$Group )
  data_b   <- rbind ( cbind ( bmi_pre, 
                              time=rep(0,len_pre )),
                      cbind ( bmi_post,
                              time=rep(1,len_post)))
  data_bmi <- data_b %>% mutate(
              time = recode(time, "0" = "baseline", 
                                  "1" = "post-measurement"))
  

# merge DLD_PRE and DLD_POST into one column
 a1 <- dplyr::select(fixed_data, c(Group,LDL_PRE ))
 b1 <- dplyr::select(fixed_data, c(Group,LDL_POST))
 ldl_pre  <- rename ( a1,LDL = LDL_PRE )
 ldl_post <- rename ( b1,LDL = LDL_POST)
 
# add a new variable that helps to illustrate time status of DLD, either baseline or  post-measurement
 len_pre  <- length( ldl_pre$Group  )
 len_post <- length( ldl_pre$Group  )
 data_l   <- rbind ( cbind(ldl_pre,time=rep(0,len_pre)),
                     cbind(ldl_post,time=rep(1,len_post)))
 data_ldl <- data_l %>% mutate(
             time = recode(time, "0" = "baseline", 
                                 "1" = "post-measurement"))

```

```{r b)ii }
  B<-tibble(  group  = pull(data_bmi,Group),
              bmi    = pull(data_bmi,BMI) ,
              time   = pull(data_bmi,time)) 
  L<-tibble(  group  = pull(data_ldl,Group),
              ldl    = pull(data_ldl,LDL) ,
              time   = pull(data_ldl,time)) 
 
   ggplot(B,aes(x    = group, 
                       y    = bmi, 
                       fill = time)) +
   geom_boxplot() +
   scale_fill_manual(values = c("blue", "red"), 
                     labels = expression("baseline", 
                                         "post-measurement")) + 
   xlab("") + 
   ylab("BMI ") + 
   theme(legend.position = "top", legend.title = element_blank())
   
   
    
   ggplot(L,aes(x    = group, 
                       y    = ldl, 
                       fill = time)) +
   geom_boxplot() +
   scale_fill_manual(values = c("blue", "red"), 
                     labels = expression("baseline", 
                                         "post-measurement")) + 
   xlab("") + 
   ylab("BMI ") + 
   theme(legend.position = "top", legend.title = element_blank())
```
